Community-Conditioned Language Models

### 2020-11-22 paper/paper.tex
scale indiscernibility by number of communities

\newcommand\Ppl{\mathsf{Ppl}}
\newcommand\IG{\mathsf{IG}}
\newcommand\Ind{I}
\DeclareMathOperator*{\avg}{avg}
\newcommand\BibTeX{B\textsc{ib}\TeX}

@@

@@

The perplexity of a CCLM gives us a measure of linguistic complexity,
but it is not a measure of how linguistically \emph{specific} the
community is. However, the CCLM induces a latent classifier,
which we call a language model-based community classifier
(LMCC). Using the Bayes theorem, we can turn a
CCLM into a LMCC: the probability that a given message $m$
belongs to a community $c_i$ is calculated as follows:
\[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m | c_i)$ (the probability that a message $m$
belongs in a community $c_i$. ) is estimated by the CCLM. $P(m)$ is the absolute
probability of a message, which can be computed as
\[P(m) = \frac 1 {\sum_i P(c=c_i) P(m|c=c_i ) }\]

We can characterise the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given (test set) message $m$ 
from community $c_j$.
\[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\]
As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).

@@

However, we want to measure the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use a perplexity-like metric,
which is has stronger information-theoretic grounds.
We define the linguistic indiscernibility of community $c_j$, noted $\Ind_j$, 
as the exponential of the entropy of the $j$th row in \(C\),
(which is a probability distribution
of predicted community for a random message from $c_j$):
\[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
Intuititely, $\Ind_j$ is a measure of the uncertainty of the LMCC,
given messages from a community $c_j$.

@@

However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1. However the r is still very weakly positive for every model.}
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

@@

\begin{table}
  \centering
  \input{floats/cclm_lmcc_ppl.tex}
  \caption{
  Pearson's $r$ correlation coefficient between community-wise linguistic perplexity and indescernibility for each model.
  }
  \label{tab:cclm_lmcc_ppl}
\end{table}

@@

On the other hand, community indiscernibility is correlated with
the average information gain of the CCLM over the baseline model,
for messages in its test set \cref{fig:comm-pca}.
Communities for which the language model bennifits most from 
the conditioned model are generally easier for the LMCC to classify.

@@

\begin{table}
  \centering
  \input{floats/lmcc_gain.tex}
  \caption{
    Pearson's \(r\) correlation coefficient between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
  }
  \label{tab:lmcc_gain}
\end{table}

>>>>

\newcommand\Ppl{\mathsf{Ppl}}
\newcommand\IG{\mathsf{IG}}
\newcommand\Ind{\mathsf{Ind}}
\DeclareMathOperator*{\avg}{average}
\newcommand\BibTeX{B\textsc{ib}\TeX}

@@

In the remainder of the paper we use $M$ to refer to messages in the test set,
and $M_j$ for the partition of the test set originating from community $c_j$.

@@

The perplexity of a CCLM gives us a measure of linguistic complexity,
but it is not a measure of how linguistically \emph{specific} the
community is. However, the CCLM induces a latent classifier,
which we call a language model-based community classifier
(LMCC). Using the Bayes theorem, we can turn a
CCLM into a LMCC: the probability that a given message $m$
belongs to a community $c_i$ is calculated as follows:
\[P(c=c_i \mid m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m \mid c_i)$ (the probability of a message $m$
in the context of community $c_i$) is estimated by the CCLM. 
$P(m)$ is the absolute
probability of a message, which can be computed as
\[P(m) = \sum_i P(c=c_i) P(m\mid c=c_i ). \]
Since the test set is balanced across communities,
we have an even prior for $P(c=c_i)$, giving us
\[P(c=c_i\mid m) \propto  P(m\mid c_i).\]

We can characterise the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
from community $c_j$.
\[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).

@@

However, we want to measure the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use a perplexity-like metric,
which is has stronger information-theoretic grounds.
To this end, we define the \emph{linguistic indiscernibility} of community $c_j$, 
noted $\Ind_j$. 
First, note that each row of $C$ is a probability distribution,
since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score.
To compute $\Ind_j$, we scale this perplexity by the number of communities, 
which makes $\Ind_j$ comparable across models trained on different
numbers of communities.\footnote{Although we only consider one set of \num{46} communities in this paper.}
\[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
Intuitively, $\Ind_j$ is a measure of the uncertainty of the LMCC,
given messages from a community $c_j$.
It ranges from $0$ (for a completely deterministic $C_j$) 
to $1$ (if $C_j$ gives equal weight to every community).

@@

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

\begin{table}
  \centering
  \input{floats/ind_corrs.tex}
  \caption{
    Pearson's $r$ correlation coefficient between community indiscernibility ($\Ind_j$) and
    two different predictors: CCLM perplexity on messages from community $c_j$ ($\Ppl_{M_j}$)
    , and CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).
  }
  \label{tab:ind-corrs}
\end{table}

@@

@@

On the other hand, community indiscernibility is correlated with
the information gain of the CCLM 
on messages from that community's test set 
(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
Communities for which the language model bennifits most from 
the conditioned model are generally easier for the LMCC to classify.

@@
